---
title: "Question 6, PSet 1"
author: "Matt Easton"
date: "2/15/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 6, PSet 1

# A) 
There is evidence of the "truncation of the extremes of socioeconomic distribution" evident in Table II, particularly when comparing the Pakistani average to the sample average of literacy and college education. The Pakistani population has an illteracy rate of 0.482, but the sample's illiteracy rate is approximately 8 percentage points lower at 0.402. This is consistent with the quoted claim above because those who are lowest in socioeconomic status are too poor to attend Hajj, and this same group also has the highest levels of illiteracy. Thus, they are underrepresented in the sample.

This same logic can explain the discrepencies in college education; the Pakistani population average is 0.21, whereas the sample is smaller at only 0.178. Those higher socioeconomic status are more likely to be college educated AND are more likely to attend Hajj through private trip planning instead of through the governmental lottery program. Thus, we can expect the sample to have a lower overall education level than the general Pakistani public.

# B) Estimate the ITT for Global Islamic Practice Index

``` {r}
#Bring in the data (downloaded .dta file from bcourses)
#install.packages("haven")
library(haven)
hajj_public <- read_dta("Downloads/hajj_public.dta")
hajj <- hajj_public

#Create an itt function
itt <- function(outcome, treatment){
  
# restrict to non-missing observations
  treatment <- treatment[!is.na(outcome)]
  outcome <- outcome[!is.na(outcome)]

# calculate itterence in means
  mean_control <- mean(outcome[treatment==0])
  mean_treated <- mean(outcome[treatment==1])
  itt <- mean_treated - mean_control
  
# prep output
  return(c(`ITT`= itt))
}

# ITT for Variable 1 x_s14aq1
itt(hajj$x_s14aq1, hajj$success)
#ITT is 0.1054

# ITT for Variable 2 x_s14aq3
itt(hajj$x_s14aq3, hajj$success)
#ITT is 0.0764

# ITT for Variable 3 x_s14aq4
itt(hajj$x_s14aq4, hajj$success)
#ITT is 0.0314

# ITT for Variable 4 x_s14aq5
itt(hajj$x_s14aq5, hajj$success)
#ITT is 0.06703

# ITT for Variable 5 x_s14aq6
itt(hajj$x_s14aq6, hajj$success)
#ITT is 0.1619

# ITT for Variable 6 x_s14bq2
itt(hajj$x_s14bq2, hajj$success)
#ITT is 0.0143

# ITT for Variable 7 x_s14aq8
itt(hajj$x_s14aq8, hajj$success)
#ITT is 0.01854

# ITT for Variable 8 x_s14aq9
itt(hajj$x_s14aq9, hajj$success)
#ITT is 0.04115

# ITT for Variable 9 x_s14aq12
itt(hajj$x_s14aq12, hajj$success)
#ITT is 0.02046

# ITT for Variable 10 x_s14aq13
itt(hajj$x_s14aq13, hajj$success)
#ITT is 0.03254

```
As we can see from the ten ITTs above, every single one is positive--indicating that winning the Hajj lottery had at least some kind of positive increase on the Global Islamic Practice among those who won the lottery.

# C) Do all of the variables in the Global Islamic Practice index belong there?
Yes, conceptually each of the ten variables for Global Islamic Practice index make sense for gauging level of commitment/practice to Islam. Perhaps the main variable that might be questionable is including whether or not the person "can read the Qu'ran". The reason this one might not make sense is because we know that almost 50% of the Pakistani population is illiterate, meaning that pratically half of the sample would automatically fail this indicator, even if they are otherwise a committed/engaged Muslim. 

# D)
While the exclusion restriction is critical to IV analysis strategy, we learned in Lecture 4 that the IVLS estimator is biased but consistent given the model. Though it is biased, it can be more or less biased based upon how large the sample size is. Therefore, we cannot assume that that this is an unbiased estimate of $\beta^k$.

# E)

