---
title: "alyssa problem set 3"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1
exercise 1
exercise 9 - (c) weight on probability of exposure; run lm with weights; weights are the inverse of the exposure probabilities; fit lm(y exposure weights = weights) ; filter for which probability10 > 0 ; calculation of weights - if exposure == 10, prob10, else prob00

y on exposure would give you the prob of spillover

all you need to find for c and (probably b) is these different probability weights based on the exposure 

exercise 11 

## Problem 5
```{r}
data <- haven :: read_dta("camp1.dta")

# use the equation for GLM model (generalized linear model)
#m <- glm(dwin tilda julyecq2 + presinc + adaaca + #I(presinc*julyecq2), data = data, family = binomial(link = "probit"))

```
b - marginal effects
get the range of the variable adaaca 
then get a sequence of values by 1 from bottom to top of range

or just get the values in the data set
get the unique values of the data set 
then get the medians of various values - median is the function 
-one of them has a median of 0 so the interaction term drops out 
-get the medians, make the data set, then make an interaction variable based on the medians of the two variables (multiply them by each other)

- you could hold all of these things and evaluate the function 
- he plotted it to look like a graph; then it looks like the graph has a constant slope

- maybe you can do analytically look it up 

c) all the unique values in the data set ; plot for the 4 scenarios;

maybe a dataset for each hypothesis
do the same for each
then joint together and graphs

adaaca as a vector of the unique values of adaaca. then repeat it 4 times as a vector. create a variable for each hypothesis. 

do predict for each, filtering for each hypothesis (type = response) because you want it on the scale of the y variable 

plot the two with facet wrap and color each hypothesis differently 

make a variable when not incumbent or when incumbent; incumbency probably matters? but he wasn't sure 
