---
title: "PS 231b PS3 q 1.1 and 2"
author: "Johanna Reyes"
date: "3/9/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1
After reading Gerber and Green Chapter 8, complete
• Exercise 1, (a)-(d) on p. 283. Note that the reference to section 8.4 in (b) should refer to section 8.2 or 8.3.

#2
Consider an experimental sample of groups with three units each. Half of all units are assigned to treatment at random. Suppose a model where there is spillover within groups but not across groups. Specifically, if any individual in a group receives treatment, the spillover effect is spread equally among members of the same group.

(a) Define the set of all potential outcomes involved in this set up.

For each group of 3 units, potential outcomes depend on whether at least one of the units in each group is assigned to treatment. The probability that each unit is assigned to treatment is 0.5, and the probability that at least 1 units in each group is assigned to treatment is 1- p(none assigned to treatment)= $1-(.5)^{3}$= 0.875.

For each group j, the potential outcome of unit i = $Y_{i_{[0,1],[0,1][0,1]}}$. $Y_{}(1,0,1)$.Each set of potential outcomes [0,1] refers to whether each unit in group j is assigned to treatment or control. 

To calculate an unbiased estimator of the treatment effect, we must weight each unit by the inverse of its probability of assignment to different treatments. We use:

$\frac

```{r}
spill_dt <- readRDS()
```


(b) Define the direct and spillover effects under the potential outcomes framework.

(c) Using the data set data_spillover.RDS provided on bCourses, estimate the direct and spillover effects.

(d) Should we use inverse propensity weights in this setting? Why / why not?